% ============================================================
% 08_roadmap.tex
% ============================================================

\section{Implementation Roadmap (2025–2030)}
\label{sec:roadmap}

\begin{table*}[t]
\centering
\scriptsize
\setlength{\tabcolsep}{5pt}
\renewcommand{\arraystretch}{1.20}
\caption{Implementation Roadmap (2025–2030): Key Milestones, Deliverables, and Impact}
\label{tab:roadmap}
\begin{tabular}{@{}p{1.9cm}p{3.5cm}p{6.9cm}p{3.1cm}@{}}
\toprule
\textbf{Year / Quarter} & \textbf{Phase} & \textbf{Key Deliverables and Milestones} & \textbf{Impact / Partners} \\
\midrule
\textbf{2025 Q4} &
\textbf{Public Launch - "Closing Breast Cancer Awareness Month with Open Data"} &
\begin{itemize}\itemsep2pt
 \item Launch \textbf{MammoChat.com} and OPTS--EGO Ledger v1.0 (Halloween).
 \item mCODE Translator API for clinical data conversion.
 \item Bilingual empathy chat (\texttt{nicegui\_chat}) pilot cohort.
 \item Register Casey DeSantis Grant deliverables on OPTS--Grant Registry.
\end{itemize} &
Lake Nona Medical City launch; AdventHealth; Florida Cancer Innovation Network \\

\textbf{2026 Q1-Q4} &
\textbf{Clinical Demonstration \& Integration} &
\begin{itemize}\itemsep2pt
 \item Raise \$12 M Series A for empathy + provenance demonstration study.
 \item HIPAA/HITECH audit and certification for OPTS--EGO.
 \item Pilots with Florida/California hospital systems.
 \item \textbf{1st Annual MammoChat Summit} (Halloween, Lake Nona)-clinicians, technologists, survivors, policy leaders.
\end{itemize} &
UCF alumni network; AdventHealth; UCSF collaborators; Patient Advocacy Council \\

\textbf{2027} &
\textbf{National Expansion \& AI Validation} &
\begin{itemize}\itemsep2pt
 \item Scale OPTS--EGO to national network.
 \item Integrate mCODE registries with NIH/ONC frameworks.
 \item AI trust audit to measure bias reduction via provenance.
 \item Publish \emph{Empathy as an Algorithm - Demonstration Outcomes}.
\end{itemize} &
NIH/NCI; ONC; National Payor Partners \\

\textbf{2028} &
\textbf{Global Partnerships \& Interoperability} &
\begin{itemize}\itemsep2pt
 \item Expand to Latin America / Europe via WHO Digital Health Ethics Network.
 \item Add Portuguese, French, Hindi.
 \item Deploy GDPR-compliant international nodes.
 \item Licensing framework for nonprofit / academic adoption.
\end{itemize} &
WHO; European Commission; Latin American Cancer Networks \\

\textbf{2029} &
\textbf{Series B \& Ecosystem Growth} &
\begin{itemize}\itemsep2pt
 \item Secure Series B; scale infrastructure and user base.
 \item 1 M verified users contributing anonymized data.
 \item MammoChat Developer SDK for third-party integrations.
 \item Pharma / device partnerships for patient-owned trials.
\end{itemize} &
Private Equity; Strategic Pharma Partners; Patient Equity Participants \\

\textbf{2030} &
\textbf{Global DAO for Patient-Owned AI Research} &
\begin{itemize}\itemsep2pt
 \item Transition OPTS--EGO to DAO with tokenized attribution.
 \item Shared governance for contributors.
 \item Establish MammoChat Foundation for open, equitable research.
\end{itemize} &
MammoChat Foundation; International Digital Health Partners \\
\bottomrule
\end{tabular}
\end{table*}
MammoChat's trajectory from prototype to platform follows a structured five-year plan that integrates scientific validation, policy alignment, and community growth.
Each phase advances a single unifying goal: to transform empathy, provenance, and patient data into the backbone of the precision-medicine economy.\autocite{prainsack2023book,ga4gh_policy,who_ai_ethics_2021}

\textbf{Q4 2025 - Public Launch.} Halloween marks the transition of MammoChat from research to reality.
The public release of \textbf{MammoChat.com} introduces the OPTS--EGO Ledger v1.0 for consent provenance, the mCODE Translator API for data standardization, and the bilingual empathy chat (\texttt{nicegui\_chat}).
Initial metrics focus on user trust, bilingual engagement, and reduction in digital anxiety.

\textbf{2026 - Demonstration and Clinical Integration.}
The \$12 M demonstration validates how empathy and provenance improve outcomes.
Deliverables: integrations with Florida/California health systems, third-party HIPAA/HITECH audit, and the OPTS--Grant Registry.
This phase demonstrates \textbf{proof of capability} (trust, participation, trial enrollment).

\textbf{2027 - National Expansion and AI Validation.}
The ledger scales to a multi-site network. Focus: payor partnerships for value-based validation, NIH/ONC mCODE registry integration, and AI trustworthiness audit with provenance-controlled datasets.

\textbf{2028 - Global Partnerships and Interoperability.}
International expansion (Latin America, Europe) with WHO Digital Health Ethics Network; deploy GDPR-compliant nodes; multi-language support (Portuguese, French, Hindi); licensing for academic / nonprofit use.

\textbf{2029 - Series B and Ecosystem Growth.}
Series B to scale operations; reach one million verified users; expand pharma/device partnerships; release the MammoChat Developer SDK.

\textbf{2030 - Global DAO for Patient-Owned AI Research.}
Transition to a patient-owned DAO with tokenized attribution and shared governance-extending the OPTS--EGO Ledger into a global commons for ethical AI in medicine.